---
title: 'EDA : Survey of Developers......'
author: "Nishant"
output:
  html_document:
    number_sections: false
    toc: true
    toc_depth: 6
    highlight: tango
    theme: cosmo
    smart: true
    code_folding: hide
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction


# Data Overview

## Loading Libraries
```{r message=FALSE, warning=FALSE}

library(dplyr)
library(ggplot2)
library(highcharter)
library(tidyverse)
library(skimr)
# library(qdap)
# library(tm)
# library(wordcloud)
library(Amelia)
library(mlbench)

```


## Loading Data

```{r message=FALSE, warning=FALSE}

Survey_results = read_csv("survey_results_public.csv")

Total_rec = nrow(Survey_results)

as.data.frame(names(Survey_results))

```

## Analysing Missing Values

We will plot a Missigness map which will show values which are missing in the data.
Color red will denote value presence and blue will denote the missing patches.

```{r message=FALSE, warning=FALSE, fig.height=10, fig.width= 10}

missmap(Survey_results, col = c("blue", "red"), legend = F)

```

As shown in the figure we can easily see there a lot of missing data only few colums from 
starting is showing continuous values. 


## Skiming the data

### Developers Code as hobby or not

```{r, message=FALSE, warning=FALSE}

Survey_results%>%
count(Hobby)%>%
  hchart("column", hcaes(x = Hobby, y = n/Total_rec*100))

```



### Country Wise Distribution of developers

```{r message=FALSE, warning=FALSE}

Survey_results$Country%>%
  skim()

Survey_results%>%
  group_by(Country)%>%
  summarise(n = n())%>%
  hchart("treemap", hcaes(x = Country, value = n, color = n))

```

### Country Wise Salary

```{r, message=FALSE, warning=FALSE}

df = Survey_results%>%
  filter(!is.na(Country))%>%
  filter(!is.na(ConvertedSalary))%>%
  group_by(Country)%>%
  summarise( min = min(ConvertedSalary), max = max(ConvertedSalary))%>%
  arrange(desc(max))%>%
  filter(min != 0)%>%
  filter(max != 0)

  highchart()%>%
  hc_xAxis(categories = df$Country)%>%
    hc_add_series(name = "Max Slary", data = df$max)%>%
    hc_add_series(name = "Minimum salary", data = df$min)%>%
  hc_chart(type = "column",
           options3d = list(enabled = TRUE, beta = 15, alpha = 15))



```



### Developer Student Or Not

```{r message=FALSE, warning=FALSE}

Survey_results$Student%>%
  skim()

Survey_results%>%
  count(Student)%>%
  hchart( "column", hcaes(x = Student, y = n))

```

### Employment Wise Distribution

```{r message=FALSE, warning=FALSE}

Survey_results$Employment%>%
  skim()

Survey_results%>%
  count(Employment)%>%
  hchart( "column", hcaes(x = Employment, y = n))

```


### Most Used Language

#### Wordcloud
```{r message=FALSE, warning=FALSE}
word_freq_lang = Survey_results%>%
  filter(!is.na(LanguageWorkedWith))%>%
  select(LanguageWorkedWith)%>%
  mutate(LanguageWorkedWith = str_split(LanguageWorkedWith, pattern = ";"))%>%
  unnest(LanguageWorkedWith)%>%
  group_by(LanguageWorkedWith)%>%
  summarise(Count = n())%>%
  arrange(desc(Count))%>%
  ungroup()%>%
  mutate(LanguageWorkedWith = reorder(LanguageWorkedWith, Count))


wordcloud(word_freq_lang$LanguageWorkedWith, word_freq_lang$Count, colors = 'blue')

```

#### Bar Chart


```{r message=FALSE, warning=FALSE}

highchart()%>%
  hc_title(text = paste("Bar Chart"))%>%
  hc_xAxis(categories = word_freq_lang$LanguageWorkedWith)%>%
  hc_add_series(data = word_freq_lang$Count, name = "Count", type = "bar")

```

### Most used Database

#### Wordcloud

```{r message=FALSE, warning=FALSE}

word_freq_database = Survey_results%>%
  filter(!is.na(DatabaseWorkedWith))%>%
  select(DatabaseWorkedWith)%>%
  mutate(DatabaseWorkedWith = str_split(DatabaseWorkedWith, pattern = ";"))%>%
  unnest(DatabaseWorkedWith)%>%
  group_by(DatabaseWorkedWith)%>%
  summarise(Count = n())%>%
  arrange(desc(Count))%>%
  ungroup()%>%
  mutate(DatabaseWorkedWith = reorder(DatabaseWorkedWith, Count))


wordcloud(word_freq_database$DatabaseWorkedWith, word_freq_database$Count, colors = 'blue')



```
#### Bar Chart


```{r message=FALSE, warning=FALSE}

highchart()%>%
  hc_title(text = paste("Bar Chart"))%>%
  hc_xAxis(categories = word_freq_database$DatabaseWorkedWith)%>%
  hc_add_series(data = word_freq_database$Count, name = "Count", type = "bar")

```


### Langauge Desired Next Year

```{r, message=FALSE, warning=FALSE}

word_freq_Lan_next = Survey_results%>%
  filter(!is.na(LanguageDesireNextYear))%>%
  select(LanguageDesireNextYear)%>%
  mutate(LanguageDesireNextYear = str_split(LanguageDesireNextYear, pattern = ";"))%>%
  unnest(LanguageDesireNextYear)%>%
  group_by(LanguageDesireNextYear)%>%
  summarise(Count = n())%>%
  arrange(desc(Count))%>%
  ungroup()%>%
  mutate(DatabaseWorkedWith = reorder(LanguageDesireNextYear, Count))


highchart()%>%
  hc_title(text = paste("Bar Chart"))%>%
  hc_xAxis(categories = word_freq_Lan_next$LanguageDesireNextYear)%>%
  hc_add_series(data = word_freq_Lan_next$Count, name = "Count", type = "bar")

```


#### Pyramid plot of Language worked with and Language worked with next year
```{r, message=FALSE, warning=FALSE}

library(plotrix)

mcol<-color.gradient(c(0,0,0.5,1),c(0,0,0.5,1),c(1,1,0.5,1),18)
fcol<-color.gradient(c(1,1,0.5,1),c(0.5,0.5,0.5,1),c(0.5,0.5,0.5,1),18)


t1 = sum(word_freq_lang$Count)
t2 = sum(word_freq_Lan_next$Count)

word_freq_lang$Count = word_freq_lang$Count/t1 *100
word_freq_Lan_next$Count = word_freq_Lan_next$Count/t2*100

par(mar = pyramid.plot(word_freq_lang$Count, word_freq_Lan_next$Count, labels = word_freq_lang$LanguageWorkedWith, top.labels=c("Langauage Worked With","","Language Desired Next Year"), main = "Language Worked with and will desire to work next", gap=5, show.values = F, rxcol = fcol, lxcol = mcol))
```




### DataBase Desired Next Year

```{r, message=FALSE, warning=FALSE}


word_freq_Database_next = Survey_results%>%
  filter(!is.na(DatabaseDesireNextYear))%>%
  select(DatabaseDesireNextYear)%>%
  mutate(DatabaseDesireNextYear = str_split(DatabaseDesireNextYear, pattern = ";"))%>%
  unnest(DatabaseDesireNextYear)%>%
  group_by(DatabaseDesireNextYear)%>%
  summarise(Count = n())%>%
  arrange(desc(Count))%>%
  ungroup()%>%
  mutate(DatabaseDesireNextYear = reorder(DatabaseDesireNextYear, Count))


highchart()%>%
  hc_title(text = paste("Bar Chart"))%>%
  hc_xAxis(categories = word_freq_Database_next$DatabaseDesireNextYear)%>%
  hc_add_series(data = word_freq_Database_next$Count, name = "Count", type = "bar")


```


#### Pyramid plot of Database worked with and Database worked with next year

```{r, message=FALSE, warning=FALSE}

library(plotrix)

mcol<-color.gradient(c(0,0,0.5,1),c(0,0,0.5,1),c(1,1,0.5,1),18)
fcol<-color.gradient(c(1,1,0.5,1),c(0.5,0.5,0.5,1),c(0.5,0.5,0.5,1),18)

t1 = sum(word_freq_database$Count)
t2 = sum(word_freq_Database_next$Count)

word_freq_database$Count = word_freq_database$Count/t1 *100
word_freq_Database_next$Count = word_freq_Database_next$Count/t2*100

par(mar = pyramid.plot(word_freq_database$Count, word_freq_Database_next$Count, labels = word_freq_database$DatabaseWorkedWith, top.labels=c("Database Worked","","Database Desired Next Year"), main = "Database Worked with and will desire to work next", gap=5, show.values = F, rxcol = fcol, lxcol = mcol))


```


### AI Interesting, Dangerous and Future Comments

```{r, message=FALSE, warning=FALSE}


Survey_results%>%
  count(AIInteresting)%>%
  hchart( "column", hcaes(x = AIInteresting, y = n))

```


```{r, message=FALSE, warning=FALSE}


Survey_results%>%
  count(AIDangerous)%>%
  hchart( "column", hcaes(x = AIDangerous, y = n))

```


```{r, message=FALSE, warning=FALSE}


Survey_results%>%
  count(AIFuture)%>%
  hchart( "column", hcaes(x = AIFuture, y = n))

```



