---
title: "KIVA: Explorartory data analysis"
author: "Nishant"
output:
  html_document:
    number_sections: false
    toc: true
    toc_depth: 6
    highlight: tango
    theme: cosmo
    smart: true
    code_folding: hide
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}

library(ggplot2)
library(dplyr)
library(lubridate)
library(qdap)
library(tm)
library(plotly)
library(lubridate)
library(DT)

Loan = read.csv("kiva_loans.csv")
location = read.csv("kiva_mpi_region_locations.csv", stringsAsFactors = F)
theme = read.csv("loan_theme_ids.csv", stringsAsFactors = F)
theme_region = read.csv("loan_themes_by_region.csv", stringsAsFactors = F)

```

## Overview of Loan data

We will overview the loan data and create a summary by loan dataframe to better analyse the data on country basis.
```{r Cleaning data, echo=FALSE}

summary_by_country = Loan %>%
  group_by(country)%>%
  summarise(no_loan = n(), LoanAmount = sum(loan_amount))

summary_by_country = summary_by_country%>%
  arrange(no_loan)

summary(summary_by_country)

```

## Analysing data grapically with each quartiles

```{r Correlation plot, message=FALSE, warning=FALSE}
first_q =  summary_by_country%>%
  filter(no_loan <= 363)%>%
ggplot(aes(x = no_loan, y = LoanAmount, col = country, size = 10, alpha = 0.4))+
  geom_point(show.legend = F)+
  scale_y_discrete("Loan Amount")+
  labs(title = "First Quartile")+
  theme(axis.ticks.y = element_blank())+
  theme_classic()

first = ggplotly(first_q)
first

second_q = summary_by_country%>%
  filter(no_loan <= 2313)%>%
  filter(no_loan > 363)%>%
  ggplot(aes(x = no_loan, y = LoanAmount, col = country, size = 10, alpha = 0.4))+
  geom_point(show.legend = F)+
  scale_y_discrete("Loan Amount")+
  labs(title = "Second Quartile")+
  theme(axis.ticks.y = element_blank())+
  theme_classic()

second = ggplotly(second_q)
second

third_q = summary_by_country%>%
  filter(no_loan <= 6687)%>%
  filter(no_loan > 2313)%>%
  ggplot(aes(x = no_loan, y = LoanAmount, col = country, size = 10, alpha = 0.4))+
  geom_point(show.legend = F)+
  scale_y_discrete("Loan Amount")+
  labs(title = "Third Quartile")+
  theme(axis.ticks.y = element_blank())+
  theme_classic()

third = ggplotly(third_q)
third

fourth_q = summary_by_country%>%
  filter(no_loan > 6687)%>%
  ggplot(aes(x = no_loan, y = LoanAmount, col = country, size = 10, alpha = 0.4))+
  geom_point(show.legend = F)+
  scale_y_discrete("Loan Amount")+
  labs(title = "Fourth Quartile")+
  theme(axis.ticks.y = element_blank())+
  theme_classic()

fourth = ggplotly(fourth_q)
fourth


```



## Analysing Data of top Countries 
Lets take a sample of countries with highest loan amount.
Initially we fo for a sample size of `10` countries.




```{r, message=FALSE, warning=FALSE}
library(zoo)

Top_10_countries = tail(summary_by_country, 10)

Loan_10 = filter(Loan, country %in% Top_10_countries$country)

Loan_10 = Loan_10 %>%
  mutate(Month = month(date), year = year(date))

Loan_10$Year_mon = with(Loan_10, sprintf("%d-%02d", year, Month))

Loan_10_summary = Loan_10%>%
  group_by(country)%>%
  summarise(Loan_amount = sum(loan_amount)/100000000)

kaggle_country = read.csv("Kaggle.csv", stringsAsFactors = F)
  
kaggle_country = kaggle_country%>%
  filter(Id %in% Loan_10$country)

kaggle_country = kaggle_country[, 1:2]

names(kaggle_country) = c('country', 'HDI')


data = left_join(Loan_10_summary, kaggle_country, by = 'country')

  ggplot(data, aes(x = as.factor(country))) +
  geom_bar(aes( y = Loan_amount, fill = 'HDI'), stat = "identity") + 
  geom_bar(aes(y = HDI * (-1), fill = 'Loan Amount'), stat = "identity")+
  scale_y_continuous(breaks = seq(-1.5, 1.5, 0.5), 
                     labels = (c(seq(1.5, 0, -0.5), seq(0.5, 1.5, 0.5))))+
  coord_flip()+
   xlab("Country")+
   ylab(" ")+
   theme(legend.title = element_blank())


 
```








```{r, message=FALSE, warning=FALSE}

library(stringr)


gender = str_split(Loan_10$borrower_genders, ",")

dt = sapply(gender, '[', 1)

Loan_10$gender = dt

Loan_10 = Loan_10 %>%
  filter(gender != "")
  

```

### Uganda

```{r, message=FALSE, warning=FALSE}

data = subset(Loan_10, country == "Uganda")

plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
 
```


```{r message=FALSE, warning=FALSE}

plot_ly(data, x = ~loan_amount, y = ~sector)%>%
  add_boxplot(color = ~sector)%>%
  layout(yaxis = list(title = " "), xaxis = list(title = "Loan Amount"), margin = list(l = 100))%>%
  rangeslider()

```


### Tajikistan

```{r, message=FALSE, warning=FALSE}

     data = subset(Loan_10, country == "Tajikistan")
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```


```{r message=FALSE, warning=FALSE}
plot_ly(data, x = ~loan_amount, y = ~sector)%>%
  add_boxplot(color = ~sector)%>%
  layout(yaxis = list(title = " "), xaxis = list(title = "Loan Amount"), margin = list(l = 100))%>%
  rangeslider()

```


### Philippines

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Philippines")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```


```{r message=FALSE, warning=FALSE}

plot_ly(data, x = ~loan_amount, y = ~sector)%>%
  add_boxplot(color = ~sector)%>%
  layout(yaxis = list(title = " "), xaxis = list(title = "Loan Amount"), margin = list(l = 100))%>%
  rangeslider()

```


### Peru

```{r, message=FALSE, warning=FALSE}

   data = subset(Loan_10, country == "Peru")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```


```{r message=FALSE, warning=FALSE}

plot_ly(data, x = ~loan_amount, y = ~sector)%>%
  add_boxplot(color = ~sector)%>%
  layout(yaxis = list(title = " "), xaxis = list(title = "Loan Amount"), margin = list(l = 100))%>%
  rangeslider()

```

### Pakistan

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Pakistan")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")

```

### Kenya

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Kenya")
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```

### El Salvador

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "El Salvador")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
   
```

### Ecuador

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Ecuador")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```


### Colombia

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Colombia")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```

### Cambodia

```{r, message=FALSE, warning=FALSE}

  data = subset(Loan_10, country == "Cambodia")
    
plot_ly(data, x = ~ repayment_interval, y = ~year, color = ~gender)%>%
  add_bars()%>%
  layout(barmode = "stack")
  
```


## Text Analysing

We will analyse activity text to get a overveiw for the general purpose of the loan. 

```{r, message=FALSE, warning=FALSE}
## Analysing data of text like sector activity and use
library(qdap)
library(tm)
library(wordcloud)

# Creating clean corpus
activity_text = VectorSource(Loan$activity)

activity_text = VCorpus(activity_text)

tm_map(activity_text, removeNumbers)
tm_map(activity_text, removePunctuation)

activity_text_tdm = TermDocumentMatrix(activity_text)

activity_text_matrix = as.matrix(activity_text_tdm)

freq = rowSums(activity_text_matrix)

#Word fequency for plotting data

word_freq = data.frame(term = names(freq), num = freq)

wordcloud(word_freq$term, word_freq$num, colors = 'red')

```

## Analysing Philippines Loan Data

When we looked into the fourth quartile the `phillippines` stand clearly as a outlier so lets take a deeper look into the data of `phillippines`.

```{r, message=FALSE, warning=FALSE}

Loan_PH = Loan%>%
  filter(country_code == "PH")

summary(Loan_PH$loan_amount)

p =  Loan_PH%>%
  ggplot(aes(sector, loan_amount))+
  scale_y_discrete("Loan Amount")+
  theme_classic()+
  stat_summary(fun.y = mean, geom = "bar", fill = "grey50")+
  theme(axis.ticks = element_blank(), axis.text.x = element_text(angle = 45, hjust =1))+
  xlab(" ")

p = ggplotly(p)

p


plot_ly(Loan_PH, x = ~date, y = ~loan_amount)
  
```

# Sector Wise Analysis(Phillippines)
Lets deep dive into the data.

```{r, message=FALSE, warning=FALSE}


Loan_PH%>%
  select(loan_amount, date, sector)%>%
  ggplot(aes(loan_amount, date))+
  geom_point()+
  facet_wrap(~ sector)+
  scale_y_discrete(breaks = NULL)+
  scale_x_discrete(breaks = NULL)+
  theme(axis.ticks.x = element_blank(), axis.ticks.y = element_blank())

```

Graphically we can identify few outliers in few sectors and if compared with previous bar plot then can see the density distribution of loan amount in few sectors.

Like looking at wholesale sector we can easily see that the amount is highest but the distribution is done in larger chunks.

**Lets look at the data of wholesale sector.**

```{r, message=FALSE, warning=FALSE}

Loan_PH_Wholesale = filter(Loan_PH, sector == "Wholesale")%>%
  mutate(year = year(date))%>%
  ggplot(aes(x = loan_amount, y = year))+
  geom_point()+
  theme_classic()

q = ggplotly(Loan_PH_Wholesale)

q


```



```{r, message=FALSE, warning=FALSE}
library(plotly)

plot_ly(Loan_10, x = Loan_10$country, type = 'histogram')

```


